Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-16 paper/v2/paper-edited.tex
5, 6

\subsection{Failure Safety}

@@

The problem of determining local preferences for each router is closely reflected in the structure of the PGIR. In particular, whenever a given router appears in multiple different contexts in the PGIR (i.e., as multiple PGIR nodes), then the compiler must decide which context to prefer.
%
In our compilation example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an advertisement from $E$ in context $(C,-,2)$ or it can receive an advertisement from its neighbor $D$ in context $(C,3,2)$.
%
The compiler's task is to determine a \emph{total ordering} of contexts for each router in the topology, which reflects the relative preference of advertisements received in each context. For example, if the compiler determines that, for router $C$, context $(C,3,2)$ is preferred to context $(C,-,2)$, written as $(C,3,2) <_{pref} (C,-,2)$, then this means that $C$ will prefer advertisements heard from $D$ (with tag $(2,2)$) over $E$ (with tag $(-,2)$). Section~\ref{sec:abgp} discusses our encoding of preferences and filters in more detail.

@@

A natural question then is how the compiler should order PGIR contexts for each router in a way that is both \emph{sound} and \emph{complete} under all possible failures? To this end, we introduce the notion of \emph{regret-free} preferences, which is motivated by the observations from the example in Figure~\ref{fig:unimplementable}. Router $r$ has a \emph{regret-free} preference for a set of advertisements $X$ over $Y$ if, whevever $r$ selects an advertisement from $X$ over another from $Y$, there is always some policy-compliant path to each final (advertisement) destination $d$ that is at least as good $(\leq_{rank})$ as any possible path to $d$ if it had accepted an advertisement from $Y$ instead. In other words, the preference of $X$ over $Y$ at $r$ is regret-free if we are never worse off by choosing from $X$.
%
In the compilation example from Figure~\ref{fig:example-compilation}, the choice for $C$ to prefer context $(C,3,2)$ to $(C,-,2)$ is regret-free, since there will always be at least as good a path to destination $W$ regardless of any failures that might occur in the network.

@@

Searching for \emph{regret-free} preferences in general is hard, and clearly enumerating all possible combinations of failures and preference orderings is intractable. We thus adopt a conservative analysis that we found to be both effective and efficient in practice. The idea is to search for regret-free preferences by comparing the set of paths available after accepting advertisments in two different PGIR contexts $N_1$ and $N_2$ while refining the comparison when necessary by taking into account information about where advertisements must have already traversed the network before arriving at $N_1$ or $N_2$.

\begin{algorithm}[t!]
\caption{Failure Protection}
\label{alg:failures}
\begin{algorithmic}[1]
\Procedure{Regret-free($G$, $N_1$, $N_2$)}{}
  \If {$loc(N_1) \neq loc(N_2)$} \Return false
  \EndIf
  \State $q \gets Queue()$
  \State $q.Enqueue (N_1, N_2)$
  \While {$!q.Empty()$}
    \State $(n_1,n_2) \gets q.Dequeue()$
    \If {$n_1 \nleq_{rank} n_2$}
      \Return false
    \EndIf
    \For {$x$ in adj($G$, $n_2$)}
      \If {\big($\exists y \in$ adj($G$,$n_1$), shadows $x$\big) \textbf{or} \\
            \hspace{5.2em} \big($\exists y \in$ $G$, dominates $n_1$, shadows $x$\big)}
        \If {$(x,y)$ not marked}
          \State {mark $(x,y)$ as seen}
          \State $q.Enqueue(x,y)$
        \EndIf
      \Else { \Return $false$}
      \EndIf
    \EndFor
  \EndWhile
  \Return true
  \EndProcedure
\end{algorithmic}
\end{algorithm}

@@

\subsection{Additional Safety Checks}

\para{Aggregation Safety}
\label{sec:agg-safety}

@@

As demonstrated with Example 2 in \S\ref{sec:motivation}, BGP aggregation can lead to subtle black-holing of traffic when failures occur. Deciding when this can happen requires knowledge of the routing policy and not just the topology. For instance, a policy might require all traffic for a particular prefix to go over a single link before being aggregated, even if there are several available in the toplogy. If that link fails, a black hole might be introduced.
%
Because the \sysname compiler has the complete user policy and topology encoded in the PGIR, it is able to efficiently check that aggregates do not black hole traffic for up to $k$ failures.

We frame the aggregation problem as a problem of connectivity in the PGIR. Specifically, for each prefix that falls under an aggregate, we find a lower bound on the number of failures required to disconnect the prefix origin from its aggregation point. The difficulty is that the same links in the topology can appear in multiple places in the PGIR.
%
%
We adopt the following simple strategy to find a lower bound on
the number of failures: $i)$ Pick a random start-to-end path in PGIR, $ii)$ remove all edges in the PGIR for the set of topological edges chosen, and $iii)$ repeat until no such path exists.
The number of PGIR paths that we are able to remove is a lower bound on the number of failures required to disconnect the prefix from its aggregate.

For example, recall the data center example from \S\ref{sec:motivation}, with the policy:
\CD{PG1} \Path \text{ }\End(\CD{A}), where \CD{PG1} falls under the \CD{PG} aggregate. Figure~\ref{fig:aggregation-safety} shows the simplified PGIR. Since the compiler knows aggregation will occur at $X$, and it knows that the \CD{PG1} prefix will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$. We could remove the $A$--$D$--$X$ path first. We would then need to remove any other $A$--$D$ or $D$--$X$ links from the PGIR (in this case none). Next, we could remove the links along the $A$--$C$--$X$ path, repeating the process. Because $A$ is then disconnected from $X$, the compiler knowns that 2 is a lower bound on the number of failures for aggregation safety for prefix \CD{PG1}. This process is repeated for other aggregation locations (e.g., $Y$).

\para{Policy Safety}

Even when programming the network centrally, it is possible for operators to make mistakes and write incorrect policies.
\sysname includes many additional safety analysis checks to try to catch common mistakes at compile time. These include: (i) preference analysis to determine when backup paths/preferences will never be used, (ii) reachability analysis to check if topology locations that should be reachable according to the operator policy are not reachable after combining both the topology and policy, (iii) anycast analysis to find instances where the operator accidentally anycasts a prefix, (iv) aggregate analysis to find when aggregates do not summarize any specific prefix, and others.

@@

Since the compiler is only able to control community tagging only for routers under the control of the AS being programmed, it cannot match on communities for external ASes. Instead, it translates matches from external ASes into a BGP regular expression filter. For example, node $D$ in Figure~\ref{fig:example-compilation} would match the single hop external paths $X$ or $Y$. In general, if routes are allowed from beyond $X$ or $Y$, these will also be captured in the BGP regular expression filters. The unknown AS topology is modelled as a special node in the PGIR that generates a filter to match any sequence of ASes.

@@

We have written a compiler for \sysname that is implemented in roughly 6700 lines of F\# code. The compiler includes command-line flags for enabling or disabling the use of the BGP MED attribute, AS path prepending, anycast, the no-export community, as well as for ensuring at least k-failure safety for user-specified aggregates. Since each prefix has a separate routing policy, the \sysname compiler is able to compile each routing policy in parallel. Currently \sysname supports generating Quagga router configurations out of the box. Users can also easily add new vendor-specific adaptors to translate from ABGP to other router configurations, or to incorporate the compiler into an existing template-based system, for example by mixing the \sysname configuration with other configuration components.

\para{PGIR construction}

Constructing automata for extended regular expressions (regular expressions including negation and intersection) are known to have high complexity~\cite{regex-complexity}. The \sysname compiler uses regular expressions derivatives~\cite{regex-derivatives} with character classes to construct deterministic automata for extended regular expressions over large alphabets efficiently. Since regular expressions are defined over a finite alphabet, and since much of the AS topology is unknown, we set the alphabet to include all uniquely referenced external ASes in the policy. Similarly, to model the unknown external AS topology beyond immediate peers, we include a special topology node to represent any unknown location.
%
Rather than construct the product graph in full, our implementation prevents exploring parts of the graph when no accepting state is reachable in any of the corresponding regular automata during construction.

@@

\para{Failure Safety}

When computing local preferences and failure safety, as described in Section~\ref{sec:compilation}, the compiler performs memoization of the \textit{protect} function. That is, whenever for two states $N_1$ and $N_2$ we compute $protect(G, N_1, N_2)$, if the function evaluates to $true$, then each of the intermediate related states $x$ and $y$ must also satisfy $protect(G, x, y)$. Memoizing these states dramatically reduces the amount of work performed to find preferences in the common case.

\para{Configuration generation}

The naive code generation algorithm described in Section~\ref{sec:compilation} is extremely memory inefficient since it generates a separate match-export pair for every unique in edge/out edge pair for every node in the product graph before minimization. Our implementation performs partial minimization during generation by recognizing common cases like when there is no restriction on exporting to or importing from neighbors.

>>>>

\subsection{Failure-Safety Analysis}

@@

The problem of determining local preferences for each router is closely reflected in the structure of the PGIR. In particular, whenever a given router appears in multiple different contexts in the PGIR (i.e., as multiple PGIR nodes that shadow each other), then the compiler must decide which context to prefer.
%
In our compilation example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an advertisement from $E$ in context $(C,-,2)$ or it can receive an advertisement from its neighbor $D$ in context $(C,3,2)$.
%
The compiler's task is to determine a \emph{total ordering} of contexts for each router in the topology, which reflects the relative preference of advertisements received in each context. For example, if the compiler determines that, for router $C$, context $(C,3,2)$ is preferred to context $(C,-,2)$, written as $(C,3,2) <_{pref} (C,-,2)$, then this means that $C$ will prefer advertisements heard from $D$ (with tag $(2,2)$) over $E$ (with tag $(-,2)$). Section~\ref{sec:abgp} discusses our encoding of preferences and filters in more detail.

@@

A natural question then is how the compiler should order PGIR contexts for each router in a way that is both \emph{sound} and \emph{complete} under all possible failures? To this end, we introduce the notion of \emph{regret-free} preferences, motivated by the observations from the example in Figure~\ref{fig:unimplementable}. Router $r$ has a \emph{regret-free} preference for a set of advertisements $X$ over $Y$ if, whevever $r$ selects an advertisement from $X$ over another from $Y$, there is always some policy-compliant path to each final (advertisement) destination $d$ that is at least as good $(\leq_{rank})$ as any possible path to $d$ if it had accepted an advertisement from $Y$ instead. In other words, the preference of $X$ over $Y$ at $r$ is regret-free if we are never worse off by choosing from $X$.
%
In the compilation example from Figure~\ref{fig:example-compilation}, the choice for $C$ to prefer context $(C,3,2)$ to $(C,-,2)$ is regret-free, since there will always be at least as good a path to destination $W$ regardless of any failures that might occur in the network.

@@

Searching for \emph{regret-free} preferences in general is hard, and enumerating all possible combinations of failures and preference orderings is intractable. We thus adopt a conservative analysis that we found to be both effective and efficient in practice. The idea is to $i)$ search for regret-free preferences by comparing the set of paths available after accepting advertisments in two different PGIR contexts $N_1$ and $N_2$, and $ii)$ refine the comparison when necessary by considering where the announcements must have traversed before arriving at $N_1$ or $N_2$.

\begin{algorithm}[t!]
\caption{Inferring regret-free preferences}
\label{alg:failures}
\begin{algorithmic}[1]
\Procedure{Regret-free($G$, $N_1$, $N_2$)}{}
  \If {$loc(N_1) \neq loc(N_2)$} \Return false
  \EndIf
  \State $q \gets Queue()$
  \State $q.Enqueue (N_1, N_2)$
  \While {$!q.Empty()$}
    \State $(n_1,n_2) \gets q.Dequeue()$
    \If {$n_1 \nleq_{rank} n_2$}
      \Return false
    \EndIf
    \For {$x$ in adj($G$, $n_2$)}
      \If {\big($\exists y \in$ adj($G$,$n_1$), shadows $x$\big) \textbf{or} \\
            \hspace{5.2em} \big($\exists y \in$ $G$, dominates $n_1$, shadows $x$\big)}
        \If {$(x,y)$ not marked}
          \State {mark $(x,y)$ as seen}
          \State $q.Enqueue(x,y)$
        \EndIf
      \Else { \Return $false$}
      \EndIf
    \EndFor
  \EndWhile
  \Return true
  \EndProcedure
\end{algorithmic}
\end{algorithm}

@@

\subsection{Aggregation-Safety Analysis}

%\para{Aggregation Safety}
\label{sec:agg-safety}

@@

As shown in \S\ref{sec:motivation}, BGP aggregation can lead to subtle black-holing of traffic when failures occur. Deciding when this can happen requires knowledge of the routing policy and not just the topology. For instance, a policy might require all traffic for a particular prefix to go over a single link before being aggregated, even if there are several available in the toplogy. If that link fails, a black hole might be introduced.
%
Because the PGIR encodes the complete user policy and topology,  \sysname can efficiently check that aggregates do not black hole traffic for up to $k$ failures.

We frame the aggregation problem as a problem of connectivity in the PGIR. Specifically, for each prefix that falls under an aggregate, we find a lower bound on the number of failures required to disconnect the prefix origin from its aggregation point. The difficulty is that the same links in the topology can appear in multiple places in the PGIR.
%

We adopt the following simple strategy to find a lower bound on
the number of failures: $i)$ pick a random start-to-end path in the PGIR, $ii)$ remove all edges in the PGIR for the set of topological edges chosen, and $iii)$ repeat until no such path exists.
The number of PGIR paths that we are able to remove is a lower bound on the number of failures required to disconnect the prefix from its aggregate.

For example, recall the data center example from \S\ref{sec:motivation}, with the policy
\CD{PG1} \Path \text{ }\End(\CD{A}), where \CD{PG1} falls under the \CD{PG} aggregate. Figure~\ref{fig:aggregation-safety} shows the simplified PGIR. Since the compiler knows aggregation will occur at $X$, and it knows that the \CD{PG1} prefix will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$. We could remove the $A$--$D$--$X$ path first. We would then need to remove any other $A$--$D$ or $D$--$X$ links from the PGIR (in this case none). Next, we could remove the links along the $A$--$C$--$X$ path, repeating the process. Because $A$ is then disconnected from $X$, the compiler knowns that 2 is a lower bound on the number of failures for aggregation safety for prefix \CD{PG1}. This process is repeated for other aggregation locations (e.g., $Y$).

@@

Since the compiler can control community tagging only for routers under the control of the AS being programmed, it cannot match on communities for external ASes. Instead, it translates matches from external ASes into a BGP regular expression filter. For example, node $D$ in Figure~\ref{fig:example-compilation} would match the single hop external paths $X$ or $Y$. In general, if routes are allowed from beyond $X$ or $Y$, these will also be captured in the BGP regular expression filters. The unknown AS topology is modelled as a special node in the PGIR that generates a filter to match any sequence of ASes.

@@

Our \sysname compiler is implemented in roughly 6700 lines of F\# code. It includes command-line flags for enabling or disabling the use of the BGP MED attribute, AS path prepending, the no-export community, as well as for ensuring at least k-failure safety for aggregate prefixes. Since each prefix has a separate routing policy, we compile each routing policy in parallel. Currently, \sysname supports generating Quagga router configurations out of the box. Users can add new vendor-specific adaptors to translate from ABGP to other router configuration languages, or incorporate the compiler into an existing template-based system, e.g., by mixing the \sysname-generated BGP configuration with other, non-BGP configuration elements.

Our compiler includes the following features that improve its performance and usability.

\para{Efficient PGIR construction}

Constructing automata for extended regular expressions (i.e., those that include negation and intersection) are known to have high complexity~\cite{regex-complexity}. The \sysname compiler uses regular expressions derivatives~\cite{regex-derivatives} with character classes to construct deterministic automata for extended regular expressions over large alphabets efficiently. Since regular expressions are defined over a finite alphabet, and since much of the AS topology is unknown, we set the alphabet to include all uniquely referenced external ASes in the policy. Similarly, to model the unknown external AS topology beyond immediate peers, we include a special topology node to represent any unknown location.
%
Rather than construct the product graph in full, our implementation prevents exploring parts of the graph when no accepting state is reachable in any of the corresponding regular automata during construction.

@@

\para{Fast failure-safety analysis}

When computing local preferences and ensuring failure safety, as described in \S\ref{sec:compilation}, the compiler performs memoization of the \textit{Regret-Free} function. That is, whenever for two states $N_1$ and $N_2$ we compute $Regret-Free(G, N_1, N_2)$, if the function evaluates to $true$, then each of the intermediate related states $x$ and $y$ must also satisfy $protect(G, x, y)$. Memoizing these states dramatically reduces the amount of work performed to find preferences in the common case.

\para{Efficient configuration generation}

The naive code generation algorithm described in \S\ref{sec:compilation} is extremely memory inefficient since it generates a separate match-export pair for every unique in edge/out edge pair for every node in the product graph before minimization. Our implementation performs partial minimization during generation by recognizing common cases such as when there is no restriction on exporting to or importing from neighbors.

\para{Checking policy correctness}

Even when programming the network centrally, it is possible for operators to make mistakes and write incorrect policies.
\sysname includes many analyses to flag common mistakes at compile time. A subset includes: (i) a preference analysis to determine when backup paths/preferences will never be used, (ii) a reachability analysis to check if topology locations that should be reachable according to the operator policy are not reachable after combining both the topology and policy, (iii) an anycast analysis to find instances where the operator accidentally anycasts a prefix (i.e., originates it from multiple locations), (iv) an aggregate analysis to find when aggregates do not summarize any specific prefix.
